Revealing Transient Strain in Geodetic Data with Gaussian Process Regression

### 2017-11-11 ResponseToReviewers/ResponseToReviewers.tex

\textbf{The method is well established. The assumptions to justify the various
hypothesis on the deterministic models of the GNSS time series are
generally well supported (seasonal signal, tectonic rate, offsets).}

\textbf{1- Now, the correlation of the stochastic processes is not completely
clear. When the authors justify the model behind equation 3, wij is a
normally distributed, uncorrelated noise. In statistic, it is equal to
a WGN process. Then, the authors introduce a separate parameter (eta)
which models the temporally correlated noise. This parameter is also
following a Gaussian distribution, with a zero-mean and a covariance
matrix $C_{\eta}$ (i.e. line 10 p5). This approach is not common in
modeling GNSS time series. The standard is based on Williams 2003,
where the author established a stochastic model with a covariance
matrix as sum of identity matrix for the white noise and another
matrix which represents the coloured noise. The coloured noise
covariance matrix is defined differently (see the literature – Bock and
Melgar Reviews of Geophysics 2016, He et al., Journal of geodynamics
2017).}

I have changed my description of the data (eq. 3 in the revised draft)
so that it includes a single term, $\eta$, that encompasses correlated
and uncorrelated noise. I then break $\eta$ down into correlated and
uncorrelated components in Section 3.1 (eq. 25). The covariance matrix
for $\eta$ in eq. 26 is more consistent with the literature.

\textbf{2- First, the author should give a summary of stochastic noise
modeling in GNSS. I also think that a discussion is needed to relate
to previous models. Why using separated stochastic models? Also I
would comment on the possibility that the (low) spatial correlation
between the parameters in your deterministic model (i.e. secular
velocities) could be absorbed in the estimation of the covariance
matrix $C_{\eta}$.}

I believe the reviewer is referring to page 4-5 of the original draft,
where I first introduce $\eta$. I agree that a more thorough
discussion on the noise model is necessary, but I do not think this is
an appropriate place in the text for it. The paper is laid out so that
Section 2 contains a general description of the method for estimating
transient strain rates, and Section 3 contains details that are
specific to the particular application. There is no universally
appropriate noise model, so I decided to save a discussion on the
noise model for Section 3. This is clarified in Section 2 of the
revised draft when I say "The appropriate noise model may vary
depending on the application, and we hold off on specifying the
covariance matrix, $C_{\eta_i}$, until Section 3.1."

@@

\textbf{Now, I am confused When looking at p.8 Section 4.1, the
description of the noise models should be given as a subsection of
Section 2 p.4. I would also improve the literature review. Several
other models have also been discussed (i.e. Montillet et al., 2014
uses a fractional Brownian motion model; ARMA, ARFIMA, GGM or Band
pass noise – see Het et al., 2017).}

Again, I think that the description of the noise model should remain
in Section 3 because it is specific to the Pacific Northwest
application.

I have added a few more references in Section 3.1 of the revised draft
to emphasize that GNSS noise modeling is a well-worked area. I refer
the reader to Bock and Melgar 2016 and He et al 2017 for a more
complete review.

@@

An Gaussian process is isotropic if its covariance function,
$C(x,x')$, can be written as a function of $||x - x'||_2^2$. An
isotropic Gaussian process is also stationary, meaning that its
statistical properties are invariant to translations. We have replace
"isotropic" with "stationary" in the text because "stationary" is a
more appropriate, and perhaps more familiar, term.

\textbf{P.20 : “ Using a compact covariance function ..” Again I would just
say that the covariance matrix is sparse and that s why you have the
mathematical simplification.}

My personal preference is to leave the wording as is.

\textbf{p.20 line 33 “ computational burden” I think that refers to p.20 line
60 “prohibitive when using several years of daily GNSS”. It would
interesting to have some number. How do you quantify the computation
time to process longer and longer time series?}

@@

This is a good point that we neglected to mention in the previous
draft. We have added a paragraph to Section 2 that explains the
reasoning and implications for this assumption. The paragraph starts
with "It should be noted that we have ignored any covariances ...".

\textbf{Page 4/5 Equation (3): Is it worth discussing at this point whether
the estimates of u and n (eta) can be separated without explicit prior
knowledge (which we may not have).  For example, orbit modeling errors
on one satellite (e.g., due to an unknown yaw problem in the
spacecraft) wilt generate a spatially and temporally correlated error
in $d_{ij}$.  How does this (stochastically non-modeled) error not project
in the u estimates?}

At this point in the text, it is assumed that the statistical
properties of u and eta are known, although not yet specified. So I do
not think this would be an appropriate place to discuss a scenario in
which we do not know u or eta. I also think that it should go with
saying that the validity of our results are contingent on having a
good model for u and eta.

In Section 3.1 and 3.2, we discuss how we choose a model for u and
eta. The key piece of "prior knowledge" that allows us to discern u
from eta is that u is negligible for inland stations. This allowed us
to isolate and constrain eta first, and then we constrained u. If we
allowed eta to be spatially correlated (which we did not), then we
would have conceivably been able to describe the spatially and
temporally correlated noise from, for example, orbit modeling errors.

@@

\textbf{Page 6: Equation 12: Maybe some additional explanation is
needed here. How are these partial derivatives formed.  My
understanding the u estimates at this point correspond one-to-one with
the positions and times of the GNSS position determinations.  Did I
miss something?}

The posterior displacements, $\hat{u}$, are actually a spatially and
temporally continuous stochastic process. We can evaluated $\hat{u}$
at any position and time that we may be interested in (not just the
position and times of the data). We are also able to compute the
spatial and temporal derivatives analytically. In Section 2 of the
revised draft, we explain how the derivatives are computed in greater
detail. We hope this clarifies any confusion.

\textbf{Page 8: Partly addresses issue raised about eqn 3.  It
probably should be noted that wether conditions east of -121 deg
longitude are different to the coast so there could be possible
problems with this noise model.}

This is also a good point that we neglected to mention in the first
draft. In Section 3.1 of the revised draft, we acknowledge the
potentially dubious assumption that noise east of -121 is
representative of the entire region. We added the disclaimer saying
"We assume the noise at these inland stations is representative of the
noise at all the stations considered in this study, which is probably
a poor assumption since the inland stations are subject to distinctly
different climatic conditions".

@@

\textbf{Editor’s Review of GJI MS 17.0593 ‘‘Revealing transient
strain...’’ by Hines and Hetland This is an novel algorithm applied to
a problem of great current interest. Based on the reviewer’s comments
and my own reading, I am recommending moderate revision. The authors’
response should include a description of their changes, and also a
version of the paper on which these can be seen (just as highlights in
the text).}

\textbf{Many of the reviewers’ comments address places where the paper
could be more clearly written, and my overall comment would be the
same. The current version perhaps spends a little too much time on the
background and not enough on explaining the method and its derivation.
As shown by the references below (these include some the authors
cite), and the Dermanis paper, which I have also attached, strain
estimation from scattered data is an area in which there has been a
lot more prior work than the authors mention. I am including these
references not because the authors need to cite many or even most of
them, but only to make the point that the problem of estimating
strain, and of looking for transients, is a familiar one: for many
readers, myself among them, much more familiar than the statistical
methods applied here. So a slightly fuller discussion of these methods
(perhaps a little more background and some reminders of, eg, what a
hyperparameter is) would be welcome to many readers; and increased
clarity of presentation usually means that the paper will be used
and cited more.}

I appreciate this comment, and I have extensively revised the paper in
order to make it more readable. To list some of the major changes:

  1) I properly define a Gaussian process in Section 2. I elaborate on
     its definition to give it a more tangible meaning.

@@

  7) I moved details of the outlier detection method to the appendix
     because it is not essential to the text.

  8) I expanded the explanation of the REML method, which is used to
     constrain u and eta in Section 3.1 and 3.2.

  9) I attempt to avoid esoteric jargon such as "hyperparameter". In
     the revised draft, "hyperparameter" has been replaced with
     "parameter", which does not degrade the precision of our
     description of the method.

  10) Most of the paper was reworded / restructured to improve
     clarity. Despite the fact that very little of the original text
     is in the revised daft, there are almost no differences in the
     content being presented.

\textbf{I feel the same way about the section on detecting outliers: again, a
well-worked area. Iterative fitting and outlier detection, as an
overall strategy, ha e been used for a long time. I appreciate that
by imposing spatial and temporal conditions you can identify outliers
not otherwise obvious, but it would help to show an example of this,
rather than the current Figure 4, in which most of the identified
cases are ones that any method could easily find.}

@@

\textbf{A few specifics (numbers are page+line):}

\textbf{2+26: possibility not risk: the latter is a technical term in seismic
hazard.}

Done

\textbf{2+47-48: I do not think this description of Shen’s method (which I’ve
used and programmed) is right, since it allows the deformation
gradients to vary in space. I’ve always thought of it as a kind of
adaptive kernel smoother, somewhat like applying 2-D loess; so it is
only dependent on nearby points, and has no assumption of uniformity
anywhere. (Actually this raises a more general question: is the method
given here use local or global support?)}

@@

\textbf{3+7: could also}

Done

\textbf{3+32: redo the sentence; as it stands, you are saying that the SCEC
exercise calculates strain rates.}

The text has been clarified to say "The method described in Holt 2013
was also tested in the SCEC transient detection exercise. Holt 2013
detects transient signal from an estimate of transient strain rates,
and they estimate transient strain rates using a ..."
